---
# IMPORTANT: Change settings here, but DO NOT change the spacing AT ALL.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Local Coffee Report"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
#Author1: "Jane Doe^[__Contributions:__  \\newline _The authors would like to thank no institution for money donated to this project. Thank you sincerely._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
#Ref1: "Some Institution, Cape Town, South Africa" # First Author's Affiliation
#Email1: "nfkatzke\\@gmail.com" # First Author's Email address

#Author2: "John Smith"
#Ref2: "Some other Institution, Cape Town, South Africa"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
#RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
#abstract: |
  #Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 8, fig.height = 5, fig.pos="H", fig.pos = 'H')

# Loading packages:
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

if(!require("pacman")) install.packages("pacman")
library(pacman)

p_load(stringi, ggridges, patchwork)

# Loading functions:
list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

# Loading raw data:
raw_coffee <- read_csv("data/Coffee/Coffee.csv")
```

# Loading and cleaning data

First, I want to inspect the data to check its format and how messy it is
```{r}
raw_count <- raw_coffee |> 
    count(roast)

raw_country_counts <- raw_coffee |> 
    count(loc_country) |> 
    arrange(desc(n))

raw_origin_pairs <- raw_coffee |> 
    distinct(origin_1, origin_2) |> 
    slice_head(n = 15)
```

In order to handle uppercase column names and load the coffee data the load_data function is used:
```{r}
coffee <- load_data("data/Coffee/Coffee.csv")
```

Additional issue with the data: the dates are in (Mon-YY) format; the country column has a typo where it reads "United States and Floyd"; and there are missing values in the roast column. Therefore:
```{r}
coffee_clean <- clean_coffee_data(coffee)
```

# Cost and rating 

First, we will investigate the relationship between cost and rating by the location of the roaster. As rating is a discrete variable, we will use a boxplot.
```{r, fig.align = 'center', fig.cap = "Cost and rating."}
box <- plot_box_cost_rating(coffee_clean)

box

save_plot(box, "Cost and rating.png")
```

The plot displays an upward trend, where higher ratings cluster at a higher cost. The mid-range of ratings overlap significantly, with long tails in both directions. As the United States (n = 1332) and Taiwan (n = 549) are the countries which roast the most coffee beans, they are the predominant countries in the data. To visualise the cost distribution of these countries the following functions are run:

```{r , fig.align = 'center', fig.cap = "The US and Taiwan."}
# Density of cost
cost_density <- plot_country_density(coffee_clean, cost_per_100g, c("United States", "Taiwan")) 
cost_density

save_plot(cost_density, "The US and Taiwan.jpg")
```

As the goal should ultimately be to find good value coffee, I define a value metric as rating divided by cost per 100g. This tells the buyer how much rating she is getting per dollar spent. The function also filters coffees with implausibly low cost. 
```{r}
coffee_clean <- add_value(coffee_clean)
```

After removing the two implausible entries, the value metric now ranges from 0.7 to 81.4, with a mean of 15.4 and a median of 16.0. Coffee costs are skewed right, indicating most coffees are reasonably priced with a few expensive coffees also present in the sample. In contrast, ratings are tightly clustered, ranging between 84 and 98. 
```{r}
desc_stats <- numeric_summary(coffee_clean) |> 
    knitr::kable(caption = "Descriptive statistics")

desc_stats
```

The following table depicts the Top 10 roasters by median value. Inclusion as a candidate for best value roaster requires at least 5 coffees. Mr. Chao Coffee, El Gran Cafe, and Jackrabbit Java lead with median values around 22-25. These roasters deliver substantially more rating per dollar spent than the sample median of 16.0. 
```{r}
candidate_roasters <- best_value_roasters(coffee_clean, min_n = 5)

t10candidates <- candidate_roasters |> 
  slice_head(n = 10) |> 
  knitr::kable(digits = 2, caption = "Top 10 roasters by value (median)")

t10candidates
```

# Indicators of good coffee \label{Indicators of good coffee }

Per the practical instructions, the most important keywords that Stelenbosch students have used as indicators of good coffee are:
- sweet
- finish
- mouthfeel
- structure
- aroma
- chocolate
- toned

Therefore, I first generate a new dataset, which includes a count of these important keywords individually as well as a total column.
```{r}
coffee_keywords <- count_keywords(coffee_clean, c("mouthfeel", "aroma", "chocolate", "sweet", "finish", "toned", "structure"))
```

This new dataset is more complete, as it includes valuable information: the frequency of keywords which Stellenbosch students regard as desirable. This is valuable information, as it provides an indicator of what qualities may make a coffee in Stellenbosch a best seller. 

# Which roaster is the 'best'?

To see whether the roasters with the best value also score well on keyword indicators, I add the average keyword count per roaster to the top 10 value rankings. This indicates there a tradeoff when choosing a roaster to supply coffee. The roasters with the highest rating (e.g. Kakalove Cafe) are more expensive, and do not necessarily have a higher frequency of desirable keywords. is  Two roasters stand out in the table as having a particularly high keyword average, Mystic Monk Coffee and Creation Food Co.

```{r}
keyword_avg <- coffee_keywords |> 
    summarise(total_keywords_mean = mean(total_keywords, na.rm = TRUE), .by = roaster)

candidate_roasters <- candidate_roasters |> 
    left_join(keyword_avg, by = "roaster")

t10complete <- candidate_roasters |> 
    slice_head(n = 10) |> 
    knitr::kable(digits = 2, caption = "Top 10 roasters by value, with average keywords")

t10complete
```

The following plot depicts the trade-off when choosing a roaster to supply coffee. As high-rating-roasters are often more expensive, and do not necessarily have more desirable characteristics (as proxied by the frequency of desirable words in the reviews). The following plot displays the relationship between rating (mean) and cost (mean) for each roasters. A clear positive relationship can be seen between rating and cost. The colour corresponds to the mean frequency of desirable words that these roasters have in their coffee (per the reviews). The three roasters that most often have desirable words in their coffee's descriptions are Mystic Monk Coffee, Creation Food Co., and Corvus Coffee Roasters. Of these three, only two appeared in the top value rankings: Mystic Monk Coffee and Creation Food Co. Therefore, it is contended these are the 'best' roasters, as they provide provide great value while having a high frequency of desirable keywords. 

```{r, fig.align = 'center', fig.cap = "Roaster Scatterplot."}
roaster_counts <- coffee_clean |> 
    count(roaster) |> 
    arrange(desc(n))

roaster_stats <- summarise_numeric(coffee_keywords, roaster) |> 
    left_join(roaster_counts, by = "roaster") |> 
    filter(n >= 5) |> 
    arrange(desc(rating_mean))

roaster_scatter <- plot_roaster_scatter(roaster_stats)

roaster_scatter

save_plot(roaster_scatter, "Roaster Scatterplot.png" )
```



