---
# IMPORTANT: Change settings here, but DO NOT change the spacing AT ALL.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Baby Names"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
#Author1: "Jane Doe^[__Contributions:__  \\newline _The authors would like to thank no institution for money donated to this project. Thank you sincerely._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
#Ref1: "Some Institution, Cape Town, South Africa" # First Author's Affiliation
#Email1: "nfkatzke\\@gmail.com" # First Author's Email address

#Author2: "John Smith"
#Ref2: "Some other Institution, Cape Town, South Africa"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
#RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
#abstract: |
  #Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

# Loading packages:
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

if(!require("pacman")) install.packages("pacman")
library(pacman)

p_load(patchwork)

# Loading functions:
list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

# Loading data:
baby_names <- read_rds("data/US_Baby_names/Baby_Names_By_US_State.rds")
t100billboard <- read_rds("data/US_Baby_names/charts.rds")
hbo_titles <- read_rds("data/US_Baby_names/HBO_titles.rds")
hbo_credits <- read_rds("data/US_Baby_names/HBO_credits.rds")
```

# Baby name persistence \label{Baby name persistence }

First, I get the national names counts, or, the sum of each name across all the states for each year and gender

```{r}
national_counts <- national_name_counts(baby_names)
glimpse(national_counts)
```

Second, I rank all the names by their count (rank 1 being the most popular), within each year-gender pairing. 
```{r}
ranked_counts <- rank_names(national_counts)
glimpse(ranked_counts)
```

Third, I create a function for the top n ranked names for year-gender pairings
```{r}
top25 <- top_n_names(ranked_counts, n = 25)
```

Fourth, in order to measure persistence, I tracked each each years top 25 names (male and female), and then checked where they ranked 1, 2 and 3 years later.The Spearman rank correlation is used, where a rank close to 1 means the names stayed in the same order, and close to 0 means they changed significantly. 
```{r}
# First, I test rank_correlation()
test_corr <- rank_correlation(top25, ranked_counts, lag = 1)

# Applying rank_correlation across lags 1-3 and combining
all_correlations <- c(1, 2, 3) |>
  map_df(~ rank_correlation(top25, ranked_counts, lag = .))

glimpse(all_correlations)
```

Fifth, I plot the correlation over time.
```{r}
persistence_plot <- rank_persistence_plot(all_correlations)
persistence_plot

save_plot(persistence_plot, "persistence_plot.png")
```

The plot does indicate that the agencies suspicions may have been correct, as there does appear to be a drop-off in persistence of the top 25 names from around 1990. The following table further confirms the agencies suspicion. At each lag, the mean correlation drops from the pre- to post-1990 period. This is possibly attributable to cultural influences such as social media and streaming causing more frequent waves of popular names. This motivates the subsequent section, which investigates drivers of baby names.
```{r}
corr_table <- correlation_period_table(all_correlations)
corr_table
```

# Investigating naming spikes \label{Investigating naming spikes }

First, I check the distribution of the data in order to get a threshold which will be use to filter out low-volume names. 500 is chosen as the threshold.
```{r}
# Checking distributions
count_summary <- summary(national_counts$count)
count_summary

count_quantiles <- quantile(national_counts$count, probs = c(0.5, 0.75, 0.9, 0.95, 0.99))
count_quantiles

```

Then, I calculate the naming spikes as year-on-year percentage changes. 
```{r}
spikes <- name_spikes(national_counts)

top_spikes <- top_spikes_table(spikes)
top_spikes
```

While a number of names are likely candidates for being driven by pop-culture influence, the name Aaliyah has the 2nd largest year-on-year spike, with a percentage change of 28 140 in 1994. As the following search indicates, the pop singer Aaliyah's first entry into the Billboard top 100 chart was also in 1994. 

```{r}
aaliyah_query <- search_billboard(t100billboard, "Aaliyah") |> 
    head(20)

aaliyah_query
```

I create the following graph in order to visualise the spike in popularity with and the rise of Aaliyah on the Billboard charts.
```{r}
aaliyah <- name_spike_plot(national_counts, "Aaliyah", "F", c(1975, 2014)) /
    billboard_weeks_plot(t100billboard, "Aaliyah", c(1975, 2014))

aaliyah

save_plot(aaliyah, "aaliyah.png")
```

This indicates that pop-culture, fashionable trends, and celebrity culture can influence naming trends over time. 