---
# IMPORTANT: Change settings here, but DO NOT change the spacing AT ALL.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Default and Credit Rating Report"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
#Author1: "Jane Doe^[__Contributions:__  \\newline _The authors would like to thank no institution for money donated to this project. Thank you sincerely._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
#Ref1: "Some Institution, Cape Town, South Africa" # First Author's Affiliation
#Email1: "nfkatzke\\@gmail.com" # First Author's Email address

#Author2: "John Smith"
#Ref2: "Some other Institution, Cape Town, South Africa"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
#RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
#abstract: |
  #Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

# Loading packages:
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

if(!require("pacman")) install.packages("pacman")
library(pacman)

p_load(readxl, ggridges, usmap)

# Loading functions:
list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

# Loading raw data:
loan_credit <- read_rds("data/Loan_Cred/loan_data.rds")
dictionary <- read_excel("data/Loan_Cred/Dictionary.xlsx")
```

# Loading data and cleaning 

I first inspect the raw data to assess how and what to clean. The dataset has 145 columns, with many containing a large amount of missing values. The columns relevant to the current analysis, such as loan status, grade, and home ownership are all relatively complete. 

```{r, results='hide'}
overview <- data_overview(loan_credit) 

```

To clean the data, I select only the relevant columns. Furthermore, I create a binary outcome variable for whether the loan status is fully paid off or not. The summary statistics of the cleaned data indicate outliers are present in the data. For instance, annual income has a maximum of 9 500 000. Similarly, debt-to-income also has a hanful of extreme outliers, with a handful of pre-cleaning observations with a dti equal to 999. The cleaning drops the most extreme dti observations. 
```{r}
loan_clean <- clean_loan_data(loan_credit)

sum_stats <- numeric_summary(loan_clean) |> 
    knitr::kable(caption = "Summary statistics")

sum_stats
```

# Default rates on short-term loans

First, I filter to short-term loans, and then create a function which calculates the default rate for any group. Among short term loans (36 months), the default rate falls as home ownership improves. Renters default at 23.1%, compared to 19.9% for owners and 15.2% for those still paying a mortgage. This provides an indication that owning vs renting matters. Although, the fact that mortgage holders default less than owners is a less intuitive result. 
```{r}
loan_short <- loan_clean |> 
    filter(term == "36 months")

# Home ownership and default rates for short-term loans
short_default_home <- default_rate_by_group(loan_short, home_ownership) |>
  knitr::kable(caption = "Default rate by home ownership")

short_default_home

# Employment length and default rates for short-term loans
short_default_emp <- default_rate_by_group(loan_short, emp_length) |>
  knitr::kable(caption = "Default rate by employment length")

short_default_emp
```

```{r, fig.align = 'center', fig.cap = "Home ownership and default rates."}
# Plotting home ownership and default rates
default_rate_home_bar <- default_rate_bar(loan_short, home_ownership)

default_rate_home_bar
```

Employment length does not appear to be a meaningful indicator of default on short-term loans. Employees with 10+ years have a slightly lower default rate, although the default rates between all employment length groups is very similar. The one notable exception is the group with a missing employment length, which defaults at 27.9%.  This is possibly attributable to unemployed or self-employed individuals, who do not report their employment length. 
```{r}
missing_income <- loan_short |>
  mutate(emp_missing = is.na(emp_length)) |>
  summarise(
    n             = n(),
    mean_income   = mean(annual_inc, na.rm = TRUE),
    median_income = median(annual_inc, na.rm = TRUE),
    .by = emp_missing
  ) |> 
    knitr::kable(caption = "Income of missing employment oberservations")

missing_income
```

# State differences in short-term loan defaults

Now, I analyse differences in the defaults on short-term loans by state. There is significant variability in the average default rate across states. States such as Arkansas, Oklahoma, Louisiana, and Alabama have the highest default rates with more than 20%. The map plot confirms that there are interstate differences in short-term loan defaults. Specifically, Southern States have a higher propensity to default. The states with the lowest propensity to default are more Northern, coastal states, such as Oregon. 

```{r, , fig.align = 'center', fig.cap = "Default rate for American states."}
state_bar <- plot_state_bar(loan_short)

state_bar
```

```{r, fig.align = 'center', fig.cap = "Default rate in the United States"}
map_plot <- plot_state_map(loan_short)

map_plot
```

# What does grade predict?

The following table displays the relationship between average default rates and credit grades.  As expected, the default rate rises steadily from Grade A to Grade G. This indicates the credit grading system tracks default risk well. The previous section highlighted two states of interest, Arkansas and Oregon, representing the top and bottom of the load default rate spectrum, respectively. Therefore, I plot the average default rate for each credit rate for both these states. I add Texas as it fell in the midrange of the state loan default rate. 

```{r}
# Default rate by grade table
default_rate_by_grade <- default_rate_by_group(loan_short, grade) |>
    arrange(desc(default_rate)) |> 
    knitr::kable(caption = "Default rate by credit grade")

default_rate_by_grade
```

As expected, Arkansas generally has a higher default rate at each credit rating level, followed by Texas and Oregon. This indicates that the grade-default relationship is consistent in this three state sub-sample. The only exception is Grade G, which makes up a small amount of the observations, indicating that the prominence of Grade G in the plot may be unreliable. 

```{r, fig.align = 'center', fig.cap = "Default rate by grade across Arkansas, Oregon, and Texas."}
# Default rate by (selected) states barplot
sample_states <- default_rate_by_grade_state(loan_short, c("AR", "TX", "OR"))

sample_states
```

