---
# IMPORTANT: Change settings here, but DO NOT change the spacing AT ALL.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Netflix Report"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
#Author1: "Jane Doe^[__Contributions:__  \\newline _The authors would like to thank no institution for money donated to this project. Thank you sincerely._]"  # First Author - note the thanks message displayed as an italic footnote of first page.
#Ref1: "Some Institution, Cape Town, South Africa" # First Author's Affiliation
#Email1: "nfkatzke\\@gmail.com" # First Author's Email address

#Author2: "John Smith"
#Ref2: "Some other Institution, Cape Town, South Africa"
#Email2: "John\\@gmail.com"
#CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

#Author3: "John Doe"
#Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
#keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
#JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
#RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
#Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
#abstract: |
  #Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

# Loading packages:
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

if(!require("pacman")) install.packages("pacman")
library(pacman)

p_load(ggridges, tidytext)

# Loading functions:
list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

# Loading raw data:
titles <- read_rds("data/netflix/titles.rds")
credits <- read_rds("data/netflix/credits.rds")
movies <- read_csv("data/netflix/netflix_movies.csv")
```

# Loading data and cleaning 

First, I filtered and cleaned the title data. I filtered the data to only inclue movies. I cleaned the data to remove column formatting issues. As some of the columns had two values per observation (e.g. multiple genres), I pivoted longer to allow for genre and country based analysis. 
```{r}
# Filtering to only movies
movies_titles <- titles |> 
  filter(type == "MOVIE")

# Cleaning the formatting of the production_countries and genres columns
movies_long <- movies_titles |> 
  unnest_list_string(production_countries) |> 
  unnest_list_string(genres)
```

# Types of movies by country 

I first identify the identify the top countries. The five countries which are the most prolific producers are the United States, India, France, and Canada. 
```{r, results='hide'}
# Top countries
top_countries <- movies_long |> 
  distinct(id, production_countries) |> 
  count(production_countries, sort = TRUE) |> 
  slice_head(n = 5)

top_countries

# Top genres by country
top_genres_by_country <- movies_long |> 
  filter(production_countries %in% top_countries$production_countries) |> 
  count(production_countries, genres, sort = TRUE) |> 
  slice_max(n, n = 5, by = production_countries)
```

```{r}
# Tabulating
top_table <- top_genres_by_country |> 
  knitr::kable(caption = "Top 5 Genres by Country")
top_table
```


The results from analysis the top genres by countries illustrate heterogeneity in preferences. Comedy and drama dominate in the United States, India, and Canada. Documentation films are strong in the United States, and the UK. 
```{r, warning = FALSE, fig.align = 'center', fig.cap = "Top Generes by Country.\\label{fig:country-genre}", fig.ext = 'png', fig.height = 5, fig.width = 7}
country_genre <- genre_by_country_plot(top_genres_by_country)

country_genre

save_plot(country_genre, "country_genre.png")
```

# Distribution of ratings by genre 

Here, I visualise the distribution of ratings by genre. History, documentation, war and sport films are more commonly well-received. Horror, family, sci-fi, and action lie at the bottom range of the rating distribution. However, most genres cluster around a mean score of between 6 and 7, and are fairly widely spread. This indicates genre alone is a weak predictor of ratings.
```{r  ,warning = FALSE, fig.align = 'center', fig.cap = "IMDb Score Distribution by Genre.\\label{fig:genreridge}", fig.ext = 'png', fig.height = 5, fig.width = 7}
genre_ridge <- genre_score_ridge_plot(movies_long)

genre_ridge

save_plot(genre_ridge, "genre_ridge.png")
```

# Distribution of runtime by genre 

Here, I visualise the distribution of runtime by genre. Similar to the ratings distribution, most genres are clustered over a similar mean (just below 100 minutes) with fairly wide spreads.
```{r , warning = FALSE, fig.align = 'center', fig.cap = "Movie Runtime Distribution by Genre.\\label{fig:runtimeviolin}", fig.ext = 'png', fig.height = 5, fig.width = 7}
runtime_violin <- genre_runtime_violin_plot(movies_long)

runtime_violin

save_plot(runtime_violin, "runtime_violin.png")
```

# Text analysis 

First, I clean and filter the dataset. To avoid double counting words from movies with multiple genres, the data is reduced to one row per movie per country before tokenising descriptions.
```{r}
movies_by_country <- movies_titles |>
    unnest_list_string(production_countries)
```

Second, I look for the top words by country. Across most countries, "life", "family", and "world" are common in movie descriptions. This indicates that a strong drama/relationship focus is appealing. The UK stands out with location-specific words like "london" and "british". Similarly, France's results lean toward "french" and "paris". This suggesting more nationally-rooted storytelling is favoured in these countries. In contrast, Canada and the United states have more broad themes, such as "family", "life" and "home". 
```{r, results='hide'}
top_words <- top_words_by_country(
    movies_by_country,
    countries = top_countries$production_countries,
    top_n = 10
)

top_words
```

```{r, warning = FALSE, fig.align = 'center', fig.cap = "Top 10 Description Words by Country.\\label{fig:wordplot}", fig.ext = 'png', fig.height = 5, fig.width = 7}
word_plot <- top_words_plot(top_words)

word_plot

save_plot(word_plot, "word_plot.png")
```

South Africa's most frequent words have a slightly different tilt than the world's top producers. Notably, frequent words include "south" and "africa", indicating the important of domestic storytelling in the country. "Apartheid" is also amongst the most frequent, illustrating the importance of the countries history. Similar to the patterns observed in the rest of the world, words such as "love", "life" and "family" are also common in South Africa. 

```{r, warning = FALSE, fig.align = 'center', fig.cap = "Top Description Words South Africa.\\label{fig:rsalollipop}", fig.ext = 'png', fig.height = 5, fig.width = 7}
rsa_lollipop <- top_words_by_country(movies_by_country, countries = "ZA", top_n = 10) |> 
  top_words_lollipop(country = "ZA", top_n = 10)

rsa_lollipop
```



